\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

\usepackage{todonotes}
\usepackage[binary-units=true,parse-numbers=true,table-auto-round=true]{siunitx}
\DeclareSIUnit\flops{FLOPS}
\DeclareSIUnit\iops{IOPS}
\sisetup{detect-weight=true}
\usepackage{float}
\usepackage{booktabs}
\usepackage{csquotes}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}


\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{A Comparison of Branch Prediction Strategies}


% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{E.R.
de Haan, A.S.J.
Misdorp, L.T.J.
van Leeuwen}
\IEEEauthorblockA{Computer Engineering Laboratory\\
Delft University of Technology\\
Delft, The Netherlands\\
Email: e.r.dehaan@student.tudelft.nl, a.s.j.misdorp@student.tudelft.nl, l.t.j.vanleeuwen@student.tudelft.nl}
}




% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}




% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\input{abstract/abstract}


\IEEEpeerreviewmaketitle



\input{introduction/introduction}
%\begin{itemize}
%\item Introduce the topic and its relevance to computer architecture.
%\item State the Research Question (RQ).
%\item Briefly describe the RQ solutions to be detailed in the report.
%\item State the main conclusions of your investigation.
%\item Report organisation (one paragraph).
%\end{itemize} 


\section{Description of the Evaluated Solutions}
In order of the year the respective papers were published they will be discussed.
The proposed branch prediction schemes of each paper will be explained to serve as context for the quantitative and qualitative comparison that follows this section.
Attention will also be paid to general branch prediction concepts, especially in the older papers.

\input{twolevel/twolevel}

\input{combined/combined}
%Describe the main idea of the solution in [1] and the way it works.
Use your own words and in case really needed you can include figures from the original paper while giving proper reference.
The text has to be self contained as the reader should be able to understand the essence of the proposed approach without reading the original paper.

\input{yags/yags}

\input{neural/neural}

\input{comparison/comparison}
%Discuss in a critical way the analysed solutions.
For each of them highlight pros and cons.
Provide a qualitative (and in case possible also a quantitative) comparison of the discussed approaches.
If applicable describe the way you would try to address the problem and explain why do you think your solution could give better results.

\input{conclusion/conclusion}
%Summarise the report content and draw the conclusions of your investigation.


% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{5}

\bibitem{twolevel}
T. Y. Yeh and Y. N. Patt, \emph{Two-Level Adaptive Training Branch Prediction}. Dept. of Elect. Eng. and Comput. Sci., The Univ. of Michigan, Ann Arbor, Michigan, Proc. of the 24th annual international symp. on Microarchitecture, pp. 51–61, Jan. 1991
\bibitem{hybrid}
S. McFarling, \emph{Combining Branch Predictors}. Digital Western Research Laboratory, Palo Alto, CA, Vol. 49. Technical Report TN-36, June 1993 

\bibitem{neural}
Jiménez, Daniel A, \emph{Fast path-based neural branch prediction}, Microarchitecture, 2003. MICRO-36. Proceedings. 36th Annual IEEE/ACM International Symposium on, IEEE, pages 243--252, 2003

\bibitem{yags}
A. N. Eden and T. Mudge, \emph{The YAGS Branch Prediction Scheme}, Proceedings of the 31st Annual ACM/IEEE International Symposium on Microarchitecture, MICRO 31, pages 69--77, 1998

\end{thebibliography}




% that's all folks
\end{document}


